---
title: "PA1_template"
output:
  html_document: default
  'html_document: keep_md: true': default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This is the course project 1 for Reproducible Research course.

### Data

The data represents the information recorded by 'personal activity monitoring device' of an individual.

### Loading and preprocessing the data
```{r}
data <- read.csv("C:/Users/qib/Desktop/R Programming/repdata_data_activity/activity.csv")
summary(data)
names(data)
```

### What is mean total number of steps taken per day?
```{r}
stepsperday <- aggregate(steps ~ date, data, sum)
stepsperday
hist(stepsperday$steps, xlab = "Total Number of Steps per Day", ylab = "Number of Days", main = "Steps taken each day")
meanspd <- mean(stepsperday$steps)
meanspd
medianspd <- median(stepsperday$steps)
medianspd
```
Mean of total steps taken per day: 10766.19
Median of total steps taken per day: 10765

### What is the average daily activity pattern?
```{r}
averagedap <- aggregate(steps ~ interval, data, mean, na.rm = TRUE)
plot(steps ~ interval, averagedap, type="l")
intervalWithmaxsteps <- averagedap[which.max(averagedap$steps),]$interval
intervalWithmaxsteps
```
The 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps: 835
### Imputing missing values
```{r}
totalmissingvalues <- sum(is.na(data$steps))
totalmissingvalues

meanstepsperinterval <- function(interval){
  averagedap[averagedap$interval==interval,]$steps
}
meanstepsperinterval

datawithnoNA <- data
for(i in 1:nrow(datawithnoNA)){
  if(is.na(datawithnoNA[i,]$steps)){
    datawithnoNA[i,]$steps <- meanstepsperinterval(datawithnoNA[i,]$interval)
  }
}

totalspdwithnoNA <- aggregate(steps ~ date, datawithnoNA, sum)
hist(totalspdwithnoNA$steps, xlab = "steps", ylab = "days", main = "Total Steps per Day with no NA")
meanstepsperdayNoNA <- mean(totalspdwithnoNA$steps)
meanstepsperdayNoNA
medianstepsperdayNoNA <- median(totalspdwithnoNA$steps)
medianstepsperdayNoNA


```
Total number of missing values: 2304
The mean did not change with/without 'NA'. However, the median changed i.e. increased.


### Are there differences in activity patterns between weekdays and weekends?
```{r}
datawithnoNA$date <- as.Date(strptime(datawithnoNA$date, format="%Y-%m-%d"))
datawithnoNA$day <- weekdays(datawithnoNA$date)
for (i in 1:nrow(datawithnoNA)) {
  if (datawithnoNA[i,]$day %in% c("Saturday","Sunday")) {
    datawithnoNA[i,]$day<-"weekend"
  }
  else{
    datawithnoNA[i,]$day<-"weekday"
  }
}
stepsbyday <- aggregate(datawithnoNA$steps ~ datawithnoNA$interval + datawithnoNA$day, datawithnoNA, mean)

names(stepsbyday) <- c("interval", "day", "steps")
library(lattice)
xyplot(steps ~ interval | day, stepsbyday, type = "l", layout = c(1, 2), 
       xlab = "Interval", ylab = "Number of steps")
```